Unusual Gentry communities

### 2021-04-13 analysis/reports/submission2/appendices/s8_gentry.Rmd
reorganize branch

---
title: "Figures and results for main manuscript"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    github_document:
       df_print: kable
       toc: true
---

@@

# Datasets by S and N (Figure 1)

s4_fig1 <- ggplot(all_di, aes(x = s0, y = n0, color = Dataset)) +
  geom_point(alpha = .1) +
  geom_point(data = filter(all_di), alpha = .4) +
  theme_bw() +
  scale_color_viridis_d(end = .9) +
  ggtitle("Communities by S and N") +
  xlab("Species richness (S); note log scale") +
  ylab("Total abundance (N); note log scale") +
  scale_x_log10() +
  scale_y_log10()

plot(s4_fig1)
```

# Illustrations of 95% interval (Figure 2)

To show the 95% interval, we need to load the distribution of shape metric values from the samples from the feasible set for a few communities. See rov_metric.md.

<!-- ```{r, fig.dim = c(7,7)} -->
<!-- library(drake) -->

<!-- db <- DBI::dbConnect(RSQLite::SQLite(), here::here("analysis", "drake", "drake-cache-net.sqlite")) -->
<!-- cache <- storr::storr_dbi("datatable", "keystable", db) -->
<!-- cache$del(key = "lock", namespace = "session") -->

<!-- net_summary <- readd(all_di_summary, cache = cache) -->

<!-- net_summary <- net_summary %>% -->
<!--   mutate(log_nparts = log(gmp:::as.double.bigz(nparts))) -->

<!-- example_fs <- readd(fs_s_44_n_13360, cache = cache) -->

<!-- example_di <- readd(di_fs_s_44_n_13360, cache =cache) -->

<!-- example_fs <- example_fs %>% -->
<!--   left_join(example_di) %>% -->
<!--   left_join(net_summary) -->

<!-- example_fs2 <- readd(fs_s_13_n_315, cache = cache) -->

<!-- example_di2 <- readd(di_fs_s_13_n_315, cache =cache) -->

<!-- example_fs2 <- example_fs2 %>% -->
<!--   left_join(example_di2) %>% -->
<!--   left_join(net_summary) -->

<!-- example_fs3 <- readd(fs_s_4_n_34, cache = cache) -->

<!-- example_di3 <- readd(di_fs_s_4_n_34, cache =cache) -->

<!-- example_fs3 <- example_fs3 %>% -->
<!--   left_join(example_di3) %>% -->
<!--   left_join(net_summary) -->

<!-- breadth_plots <- list( -->

<!--   ggplot(example_fs3, aes(rank, abund, group = sim, color = skew)) + -->
<!--     geom_line(alpha = .25) + -->
<!--     theme_bw() + -->
<!--     scale_color_viridis_c(option = "plasma", end = .8) + -->
<!--     ggtitle("Small community", subtitle = paste0("S = ", (example_fs3$s0), "; N = ", (example_fs3$n0[1]))) + -->
<!--     theme(legend.position = "right") + -->
<!--     xlab("Rank") + -->
<!--     ylab("Abundance"), -->

<!--   ggplot(example_fs3, aes(skew)) + -->
<!--   #  geom_density() + -->
<!--     geom_histogram(bins = 50) + -->
<!--     theme_bw() + -->
<!--     geom_vline(xintercept = c(example_fs3$skew_97p5[1], example_fs3$skew_2p5[1]), color = "red") + -->
<!--     ggtitle("", subtitle = paste0("Breadth index: ", round((example_fs3$skew_95_ratio_2t[1]), 2))) + -->
<!--     xlab("Skewness") + -->
<!--     ylab("Count"), -->

<!--   ggplot(example_fs2, aes(rank, abund, group = sim, color = skew)) + -->
<!--     geom_line(alpha = .1) + -->
<!--     theme_bw() + -->
<!--     scale_color_viridis_c(option = "plasma", end = .8) + -->
<!--     ggtitle("Medium community", subtitle = paste0("S = ", (example_fs2$s0), "; N = ", (example_fs2$n0[1]))) + -->
<!--     theme(legend.position = "right")+ -->
<!--     xlab("Rank") + -->
<!--     ylab("Abundance") + -->
<!--     ylim(0, 200), # Remove 3 sads that make the axes too big to be interpretable -->

<!--   ggplot(example_fs2, aes(skew)) + -->
<!--   #  geom_density() + -->
<!--     geom_histogram(bins = 50) + -->
<!--     theme_bw() + -->
<!--     geom_vline(xintercept = c(example_fs2$skew_97p5[1], example_fs2$skew_2p5[1]), color = "red") + -->
<!--     ggtitle("", subtitle =  paste0("Breadth index: ", round((example_fs2$skew_95_ratio_1t[1]), 2)))+ -->
<!--     xlab("Skewness") + -->
<!--     ylab("Count"), -->

<!--   ggplot(example_fs, aes(rank, abund, group = sim, color = skew)) + -->
<!--     geom_line(alpha = .1) + -->
<!--     theme_bw() + -->
<!--     scale_color_viridis_c(option = "plasma", end = .8) + -->
<!--     ggtitle("Large community", subtitle = paste0("S = ", (example_fs$s0), "; N = ", (example_fs$n0[1]))) + -->
<!--     theme(legend.position = "right") + -->
<!--     ylim(0, 4000) + # Remove a very few very very uneven SADs that make the scale too big to be interpretable -->
<!--     theme(axis.text.y = element_text(size = 6, angle = 60))+ -->
<!--     xlab("Rank") + -->
<!--     ylab("Abundance"), -->

<!--   ggplot(example_fs, aes(skew)) + -->
<!--    # geom_density() + -->
<!--     geom_histogram(bins = 50) + -->
<!--     theme_bw() + -->
<!--     geom_vline(xintercept = c(example_fs$skew_97p5[1], example_fs$skew_2p5[1]), color = "red") + -->
<!--     ggtitle("", subtitle =  paste0("Breadth index: ", round((example_fs$skew_95_ratio_1t[1]), 2)))+ -->
<!--     xlab("Skewness") + -->
<!--     ylab("Count") -->

<!-- ) -->

<!-- fig_1 <- gridExtra::grid.arrange(grobs = breadth_plots, ncol = 2, top = textGrob("Figure 1", gp = gpar(fill = "white"))) -->

<!-- plot(fig_1) -->

<!-- DBI::dbDisconnect(db) -->
<!-- rm(cache) -->
<!-- rm(db) -->
<!-- ``` -->

# Dissimilarity (Supplement)

@@

```{r, fig.dim = c(4, 4)}
ggplot(all_di, aes(sim_pos_from_best, real_po, color = `Observed \npercentile score`)) +
  geom_point(alpha = .3) + 
  geom_point(data = filter(all_di, s0 < 0)) +
  geom_abline(slope = 1, intercept = 0) +
  scale_color_viridis_d(end = .8, direction = -1) +
  xlab("Mean of elements from feasible set") +
  ylab("Observed value") +
  ggtitle("Dissimilarity to the central tendency") +  theme(legend.position = "bottom")#,
# 
# 
# ggplot(filter(all_di, nparts < 10 ^ 50), aes(nparts, sim_pos_from_best, color = Dataset)) +
#   geom_point(alpha = .3) + 
#   geom_point(data = filter(all_di, s0 < 0)) +
#   scale_color_viridis_d(end = .9) +
#   xlab("Number of elements in the feasible set") +
#   ylab("Mean dissimilarity of feasible set to central tendency") +
#   scale_x_log10() +
#   ggtitle("Over the size of the feasible set")
# ), ncol = 1,
#top = textGrob("Dissimilarity to central tendency"))

# Metric histograms by dataset (Figure 4)

```{r first hists, fig.dim = c(10,5)}

fig_2 <- gridExtra::grid.arrange(grobs = list(

  plot_percentile_hist(all_di, "real_po_percentile_mean", "Dissimilarity to the \ncentral tendency", facetvar = "Dataset", tails = 1),
  plot_percentile_hist(all_di, "nsingletons_percentile_mean", "\nNumber of rare species", facetvar = "Dataset"),
  plot_percentile_hist(all_di, "skew_percentile_mean", "\nSkewness", facetvar = "Dataset", min_s0 = 3),
  plot_percentile_hist(all_di, "simpson_percentile_mean", "\nSimpson evenness", facetvar = "Dataset"),
  plot_percentile_hist(all_di, "shannon_percentile_mean", "\nShannon diversity", facetvar = "Dataset")
), ncol = 5,
top = textGrob("Figure 2", gp = gpar(fill = "white")), left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")),
bottom = textGrob("Percentile rank of observed value relative to feasible set"), gp = gpar(fill = "white"))
```

# Results table

```{r}
diss_results <- all_di %>%
  filter(nparts > 20) %>%
  group_by(Dataset) %>%
  summarize(high_diss = mean(real_po_percentile_excl > 95),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High dissimilarity` = paste0(signif(100 * high_diss, 2), "%; n = ", nsites)) %>%
  select(-nsites)

skew_results <- all_di %>%
  filter(nparts > 40, s0 > 2) %>%
  group_by(Dataset) %>%
  summarize(high_skew = mean(skew_percentile_excl > 97.5),
            low_skew = mean(skew_percentile < 2.5),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High skew` = paste0(signif(100 * high_skew, 2), "%; n = ", nsites),
         `Low skew` = paste0(signif(100 * low_skew, 2), "%; n = ", nsites)) %>%
  select(-nsites)

metric_results <- all_di %>%
  filter(nparts > 40) %>%
  group_by(Dataset) %>%
  summarize(high_simpson = mean(simpson_percentile_excl > 97.5),
            low_simpson = mean(simpson_percentile < 2.5),
            high_shannon = mean(shannon_percentile_excl > 97.5),
            low_shannon = mean(shannon_percentile < 2.5),
            high_nsingletons = mean(nsingletons_percentile_excl > 97.5),
            low_nsingletons = mean(nsingletons_percentile < 2.5),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High Simpson` = paste0(signif(100 * high_simpson, 2), "%; n = ", nsites),
         `Low Simpson` = paste0(signif(100 * low_simpson, 2), "%; n = ", nsites),
         `High Shannon` = paste0(signif(100 * high_shannon, 2), "%; n = ", nsites),
         `Low Shannon` = paste0(signif(100 * low_shannon, 2), "%; n = ", nsites),
         `High proportion of rare species` = paste0(signif(100 * high_nsingletons, 2), "%; n = ", nsites),
         `Low proportion of rare species` = paste0(signif(100 * low_nsingletons, 2), "%; n = ", nsites)) %>%
  select(-nsites)

all_results <- left_join(diss_results, skew_results) %>%
  left_join(metric_results)

cols1 <- c("Dataset", "High dissimilarity", "High proportion of rare species", "High skew", "Low Simpson", "Low Shannon")

all_results %>%
  select(cols1)

cols2 <- c("Dataset", "Low proportion of rare species", "Low skew", "High Simpson", "High Shannon")

all_results %>%
  select(cols2)

# Narrowness

@@

```{r, fig.dim = c(10,5)}

fig_2 <- gridExtra::grid.arrange(grobs = list(

  plot_narrowness_hist(all_di, "sim_pos_from_best", "Dissimilarity to the \ncentral tendency", facetvar = "Dataset"),
  plot_narrowness_hist(all_di, "nsingletons_95_ratio_2t", "\nNumber of rare species", facetvar = "Dataset"),
  plot_narrowness_hist(all_di, "skew_95_ratio_2t", "\nSkewness", facetvar = "Dataset", min_s0 = 3),
  plot_narrowness_hist(all_di, "simpson_95_ratio_2t", "\nSimpson evenness", facetvar = "Dataset"),
  plot_narrowness_hist(all_di, "shannon_95_ratio_2t", "\nShannon diversity", facetvar = "Dataset")
), ncol = 5,
 left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")), gp = gpar(fill = "white"))

@@

# Gentry

gentry_di <- gentry_di %>%
  mutate(low_avg_abund = ifelse(n0/s0 < 3, "N:S < 3", "N:S > 3"))

fig_2 <- gridExtra::grid.arrange(grobs = list(

@@

# Comparison of FIA and comparably sized communities (Figure 4)

```{r, fig.dim = c(10, 5)}

fia_max_s = max(filter(all_di, dat == "fia")$s0)
fia_max_n = max(filter(all_di, dat == "fia")$n0)

fia = filter(all_di, dat == "fia")

small_di <- all_di %>%
  filter(s0 <= fia_max_s,
         n0 <= fia_max_n) %>%
  mutate(fia_yn = ifelse(dat == "fia", "fia", "other datasets"),
         right_corner = FALSE)

for(i in 1:nrow(small_di)) {
  if(small_di$dat[i] != "fia") {
    fia_match_s0 = filter(fia, s0 >= small_di$s0[i])

    max_n0_for_this_s0 = max(fia_match_s0$n0)

    small_di$right_corner[i] = small_di$n0[i] > max_n0_for_this_s0

  }
}

small_di <- small_di %>%
  filter(!right_corner)
not_fia <- filter(small_di, fia_yn != "fia")
fia = filter(small_di, dat == "fia")

small_di_s_n <- small_di %>%
  select(s0, n0) %>%
  distinct() %>%
  mutate(in_fia = FALSE,
         in_not_fia = FALSE)

for(i in 1:nrow(small_di_s_n)) {

  this_s0 = small_di_s_n$s0[i]
  this_n0 = small_di_s_n$n0[i]

  fia_matches = filter(fia, s0 == this_s0, n0 == this_n0)
  not_fia_matches = filter(not_fia, s0 == this_s0, n0 == this_n0)

  small_di_s_n$in_fia[i] = nrow(fia_matches) > 0
  small_di_s_n$in_not_fia[i] = nrow(not_fia_matches) > 0

small_di_s_n <- filter(small_di_s_n, in_fia, in_not_fia)

set.seed(1977)

subsamples <- list()

for(i in 1:nrow(small_di_s_n)) {

  this_s0 = small_di_s_n$s0[i]
  this_n0 = small_di_s_n$n0[i]

  fia_matches = filter(fia, s0 == this_s0, n0 == this_n0)

  not_fia_matches = filter(not_fia, s0 == this_s0, n0 == this_n0)

  n_to_draw = min(nrow(fia_matches), nrow(not_fia_matches))

  fia_draw = sample.int(n = nrow(fia_matches), size = n_to_draw, replace = F)
  not_fia_draw = sample.int(n = nrow(not_fia_matches), size = n_to_draw, replace = F)

  subsamples[[i]] <- (bind_rows(fia_matches[fia_draw,], not_fia_matches[not_fia_draw, ]))

sub_di <- bind_rows(subsamples) %>%
  mutate(Dataset = ifelse(fia_yn == "fia", "FIA", "Other datasets"))

fia_di <- filter(sub_di, fia_yn == "fia")
other_di <- filter(sub_di, fia_yn != "fia")

ks_compare <- function(fia_df, other_df, compare_var) {

  if(grepl("skew", compare_var)) {
    fia_df <- filter(fia_df, s0 > 2)
    other_df <- filter(fia_df, s0 >2)
  }

  if(grepl("sim_pos", compare_var)) {
    fia_df <- filter(fia_df, nparts > 20)
    other_df <- filter(other_df, nparts > 20)
  } else {
      fia_df <- filter(fia_df, nparts > 40)
    other_df <- filter(other_df, nparts > 40)
  }

  kstest <- ks.test(fia_df[[compare_var]], other_df[[compare_var]])

  return(data.frame(
    var = compare_var,
    d = kstest$statistic,
    p = kstest$p.value
  ))
}

simpson_rat_ks <- ks_compare(fia_di, other_di, "simpson_95_ratio_2t")
skew_rat_ks <- ks_compare(fia_di, other_di, "skew_95_ratio_2t")
shannon_rat_ks <- ks_compare(fia_di, other_di, "shannon_95_ratio_2t")
nsingletons_rat_ks <- ks_compare(fia_di, other_di, "nsingletons_95_ratio_2t")
diss_rat_ks <- ks_compare(fia_di, other_di, "sim_pos_from_best")

bind_rows(simpson_rat_ks, skew_rat_ks, shannon_rat_ks, nsingletons_rat_ks, diss_rat_ks)

simpson_perc_ks <- ks_compare(fia_di, other_di, "simpson_percentile")
skew_perc_ks <- ks_compare(fia_di, other_di, "skew_percentile_excl")
shannon_perc_ks <- ks_compare(fia_di, other_di, "shannon_percentile")
nsingletons_perc_ks <- ks_compare(fia_di, other_di, "nsingletons_percentile_excl")
diss_perc_ks <- ks_compare(fia_di, other_di, "real_po_percentile_excl")

bind_rows(simpson_perc_ks, skew_perc_ks, shannon_perc_ks, nsingletons_perc_ks, diss_perc_ks)

fig_2 <- gridExtra::grid.arrange(grobs = list(

  plot_percentile_hist(sub_di, "real_po_percentile_mean", "Dissimilarity to the \ncentral tendency", facetvar = "Dataset", tails = 1),
  plot_percentile_hist(sub_di, "nsingletons_percentile_mean", "\nNumber of rare species", facetvar = "Dataset"),
  plot_percentile_hist(sub_di, "skew_percentile_mean", "\nSkewness", facetvar = "Dataset", min_s0 = 3),
  plot_percentile_hist(sub_di, "simpson_percentile_mean", "\nSimpson evenness", facetvar = "Dataset"),
  plot_percentile_hist(sub_di, "shannon_percentile_mean", "\nShannon diversity", facetvar = "Dataset")
), ncol = 5,
top = textGrob("Figure 2", gp = gpar(fill = "white")), left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")),
bottom = textGrob("Percentile rank of observed value relative to feasible set"), gp = gpar(fill = "white"))

fig_2 <- gridExtra::grid.arrange(grobs = list(

  plot_narrowness_hist(sub_di, "sim_pos_from_best", "Dissimilarity to the \ncentral tendency", facetvar = "Dataset"),
  plot_narrowness_hist(sub_di, "nsingletons_95_ratio_2t", "\nNumber of rare species", facetvar = "Dataset"),
  plot_narrowness_hist(sub_di, "skew_95_ratio_2t", "\nSkewness", facetvar = "Dataset", min_s0 = 3),
  plot_narrowness_hist(sub_di, "simpson_95_ratio_2t", "\nSimpson evenness", facetvar = "Dataset"),
  plot_narrowness_hist(sub_di, "shannon_95_ratio_2t", "\nShannon diversity", facetvar = "Dataset")
), ncol = 5,
 left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")), gp = gpar(fill = "white"))

diss_results <- sub_di %>%
  filter(nparts > 20) %>%
  group_by(Dataset) %>%
  summarize(high_diss = mean(real_po_percentile_excl > 95),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High dissimilarity` = paste0(signif(100 * high_diss, 2), "%; n = ", nsites)) %>%
  select(-nsites)

skew_results <- sub_di %>%
  filter(nparts > 40, s0 > 2) %>%
  group_by(Dataset) %>%
  summarize(high_skew = mean(skew_percentile_excl > 97.5),
            low_skew = mean(skew_percentile < 2.5),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High skew` = paste0(signif(100 * high_skew, 2), "%; n = ", nsites),
         `Low skew` = paste0(signif(100 * low_skew, 2), "%; n = ", nsites)) %>%
  select(-nsites)

metric_results <- sub_di %>%
  filter(nparts > 40) %>%
  group_by(Dataset) %>%
  summarize(high_simpson = mean(simpson_percentile_excl > 97.5),
            low_simpson = mean(simpson_percentile < 2.5),
            high_shannon = mean(shannon_percentile_excl > 97.5),
            low_shannon = mean(shannon_percentile < 2.5),
            high_nsingletons = mean(nsingletons_percentile_excl > 97.5),
            low_nsingletons = mean(nsingletons_percentile < 2.5),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High Simpson` = paste0(signif(100 * high_simpson, 2), "%; n = ", nsites),
         `Low Simpson` = paste0(signif(100 * low_simpson, 2), "%; n = ", nsites),
         `High Shannon` = paste0(signif(100 * high_shannon, 2), "%; n = ", nsites),
         `Low Shannon` = paste0(signif(100 * low_shannon, 2), "%; n = ", nsites),
         `High proportion of rare species` = paste0(signif(100 * high_nsingletons, 2), "%; n = ", nsites),
         `Low proportion of rare species` = paste0(signif(100 * low_nsingletons, 2), "%; n = ", nsites)) %>%
  select(-nsites)

all_results <- left_join(diss_results, skew_results) %>%
  left_join(metric_results)

cols1 <- c("Dataset", "High dissimilarity", "High proportion of rare species", "High skew", "Low Simpson", "Low Shannon")

all_results %>%
  select(cols1)

# Very small communities

all_di <- all_di %>%
  mutate( `Number of elements` = ifelse(nparts < 2000, "Less than 2000", "More than 2000"))
diss_results <- all_di %>%
  filter(nparts > 20) %>%
  group_by(Dataset,  `Number of elements` ) %>%
  summarize(high_diss = mean(real_po_percentile_excl > 95),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High dissimilarity` = paste0(signif(100 * high_diss, 2), "%; n = ", nsites)) %>%
  select(-nsites)

skew_results <- all_di %>%
  filter(nparts > 40, s0 > 2) %>%
  group_by(Dataset,  `Number of elements` ) %>%
  summarize(high_skew = mean(skew_percentile_excl > 97.5),
            low_skew = mean(skew_percentile < 2.5),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High skew` = paste0(signif(100 * high_skew, 2), "%; n = ", nsites),
         `Low skew` = paste0(signif(100 * low_skew, 2), "%; n = ", nsites)) %>%
  select(-nsites)

metric_results <- all_di %>%
  filter(nparts > 40) %>%
  group_by(Dataset,  `Number of elements` ) %>%
  summarize(high_simpson = mean(simpson_percentile_excl > 97.5),
            low_simpson = mean(simpson_percentile < 2.5),
            high_shannon = mean(shannon_percentile_excl > 97.5),
            low_shannon = mean(shannon_percentile < 2.5),
            high_nsingletons = mean(nsingletons_percentile_excl > 97.5),
            low_nsingletons = mean(nsingletons_percentile < 2.5),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High Simpson` = paste0(signif(100 * high_simpson, 2), "%; n = ", nsites),
         `Low Simpson` = paste0(signif(100 * low_simpson, 2), "%; n = ", nsites),
         `High Shannon` = paste0(signif(100 * high_shannon, 2), "%; n = ", nsites),
         `Low Shannon` = paste0(signif(100 * low_shannon, 2), "%; n = ", nsites),
         `High proportion of rare species` = paste0(signif(100 * high_nsingletons, 2), "%; n = ", nsites),
         `Low proportion of rare species` = paste0(signif(100 * low_nsingletons, 2), "%; n = ", nsites)) %>%
  select(-nsites)

all_results <- left_join(diss_results, skew_results) %>%
  left_join(metric_results)

cols1 <- c("Dataset", "Number of elements", "High dissimilarity", "High proportion of rare species", "High skew", "Low Simpson", "Low Shannon")

all_results %>%
  select(cols1)

plot_narrowness_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "sim_pos_from_best", "Dissimilarity") +
  facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)

plot_percentile_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "real_po_percentile_mean", "Dissimilarity") +
  facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)

plot_narrowness_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "nsingletons_95_ratio_2t", "Number of rare species") +
  facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)

plot_percentile_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "nsingletons_percentile_mean", "Number of rare species") +
  facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)

plot_narrowness_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "simpson_95_ratio_2t", "Simpson") +
  facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)

plot_percentile_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "simpson_percentile_mean", "Simpson") +
  facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)

plot_narrowness_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "shannon_95_ratio_2t", "Shannon") +
  facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)

plot_percentile_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "shannon_percentile_mean", "Shannon") +
  facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)

ggplot(all_di, aes(s0, n0, color = `Number of elements`)) +
  geom_point() +
  xlim(0, 30) +
  ylim(0,150)

all_di <- all_di %>%
  mutate( `Number of elements` = ifelse(nparts < 1000, "Less than 1000", "More than 1000"))
diss_results <- all_di %>%
  filter(nparts > 20) %>%
  group_by(Dataset,  `Number of elements`) %>%
  summarize(high_diss = mean(real_po_percentile_excl > 95),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High dissimilarity` = paste0(signif(100 * high_diss, 2), "%; n = ", nsites)) %>%
  select(-nsites)
ggplot(filter(all_di, dat == "mcdb"), aes(skew_95_ratio_2t)) + 
  geom_histogram() +
  facet_wrap(vars(`Number of elements`), scales = "free_y")

skew_results <- all_di %>%
  filter(nparts > 40, s0 > 2) %>%
  group_by(Dataset,  `Number of elements`) %>%
  summarize(high_skew = mean(skew_percentile_excl > 97.5),
            low_skew = mean(skew_percentile < 2.5),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High skew` = paste0(signif(100 * high_skew, 2), "%; n = ", nsites),
         `Low skew` = paste0(signif(100 * low_skew, 2), "%; n = ", nsites)) %>%
  select(-nsites)

metric_results <- all_di %>%
  filter(nparts > 40) %>%
  group_by(Dataset, `Number of elements`) %>%
  summarize(high_simpson = mean(simpson_percentile_excl > 97.5),
            low_simpson = mean(simpson_percentile < 2.5),
            high_shannon = mean(shannon_percentile_excl > 97.5),
            low_shannon = mean(shannon_percentile < 2.5),
            high_nsingletons = mean(nsingletons_percentile_excl > 97.5),
            low_nsingletons = mean(nsingletons_percentile < 2.5),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High Simpson` = paste0(signif(100 * high_simpson, 2), "%; n = ", nsites),
         `Low Simpson` = paste0(signif(100 * low_simpson, 2), "%; n = ", nsites),
         `High Shannon` = paste0(signif(100 * high_shannon, 2), "%; n = ", nsites),
         `Low Shannon` = paste0(signif(100 * low_shannon, 2), "%; n = ", nsites),
         `High proportion of rare species` = paste0(signif(100 * high_nsingletons, 2), "%; n = ", nsites),
         `Low proportion of rare species` = paste0(signif(100 * low_nsingletons, 2), "%; n = ", nsites)) %>%
  select(-nsites)

all_results <- left_join(diss_results, skew_results) %>%
  left_join(metric_results)

cols1 <- c("Dataset", "Number of elements", "High dissimilarity", "High proportion of rare species", "High skew", "Low Simpson", "Low Shannon")

all_results %>%
  select(cols1)

# Resampling

all_di_with_resamples <- read.csv(here::here("analysis", "rev_prototyping", "all_di.csv"))

jk_di_full <- read.csv(here::here("analysis", "rev_prototyping", "all_di_jk.csv"))

all_ct_with_resamples <- read.csv(here::here("analysis", "rev_prototyping", "all_ct.csv"))

jk_ct <- read.csv(here::here("analysis", "rev_prototyping", "all_ct_jk.csv"))

jk_di_full <- left_join(jk_di_full, jk_ct)

all_di_with_resamples <- left_join(all_di_with_resamples, all_ct_with_resamples)

jk_di <- jk_di_full %>%
  group_by_all() %>%
  mutate(site_source = unlist(strsplit(site, "_")[[1]][[1]])) %>%
  ungroup() %>%
  select(dat, site, site_source, s0, n0, skew_percentile_excl,skew_percentile, simpson_percentile, simpson_percentile_excl, shannon_percentile, shannon_percentile_excl, nsingletons_percentile, nsingletons_percentile_excl, real_po_percentile, real_po_percentile_excl, sim_pos_from_best, skew_95_ratio_2t, simpson_95_ratio_2t, shannon_95_ratio_2t, nparts)

skewcols <- colnames(jk_di)[ grepl("skew", colnames(jk_di))]

jk_di[ which(jk_di$s0 < 3), skewcols] <- NA

jk_di_means <- jk_di %>%
  filter(n0 != s0,
         s0 != 1,
         n0 != (s0 + 1)) %>%
  group_by(dat, site_source) %>%
  mutate(njks = dplyr::n(),
         njks_skew = sum(!is.na(skew_percentile)))%>%
  summarise_at(vars(-group_cols(), -site), mean, na.rm =  T) %>%
  mutate(resampling = "Subsampling") %>%
  mutate(site = as.numeric(site_source)) %>%
#  select(-site_source) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
         Dataset = ifelse(Dataset == "fia", "FIA", Dataset),
         Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
         Dataset = ifelse(Dataset == "mcdb", "Mammal Communities", Dataset),
         Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
         Dataset = ifelse(Dataset == "misc_abund", "Misc. Abundance", Dataset)) %>%
  filter(nparts > 20,
         njks == 10) %>%
  ungroup()

jk_di[ which(jk_di$njks_skew < 10), skewcols] <- NA

skewcols <- colnames(all_di_with_resamples[ which(grepl("skew", colnames(all_di_with_resamples)))])

all_di_with_resamples[ which(all_di_with_resamples$s0 < 3), skewcols] <- NA

all_di_with_resamples <- all_di_with_resamples %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log_s0 = log(s0),
         log_n0 = log(n0)) %>%
  filter(n0 != s0,
         s0 != 1,
       #  !singletons,
         n0 != (s0 + 1)) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
         Dataset = ifelse(Dataset == "fia", "FIA", Dataset),
         Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
         Dataset = ifelse(Dataset == "mcdb", "Mammal Communities", Dataset),
         Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
         Dataset = ifelse(Dataset == "misc_abund", "Misc. Abundance", Dataset)) %>%
  filter(nparts > 20) %>%
  mutate(resampling = ifelse(singletons, "Rare species", "Raw")) %>%
  bind_rows(jk_di_means)

metric_rs_results <- all_di_with_resamples %>%
  group_by(Dataset, resampling) %>%
  summarize(
    skew_high = mean(skew_percentile_excl > 97.5, na.rm = T),
    skew_low = mean(skew_percentile < 2.5, na.rm = T),
    even_high = mean(simpson_percentile_excl > 97.5, na.rm = T),
    even_low = mean(simpson_percentile < 2.5, na.rm = T),
    shannon_high = mean(shannon_percentile_excl > 97.5, na.rm = T),
    shannon_low = mean(shannon_percentile < 2.5, na.rm = T),
    nsingletons_high = mean(nsingletons_percentile_excl > 97.5, na.rm = T),
    nsingletons_low = mean(nsingletons_percentile < 2.5, na.rm = T),
    nskew = sum(!is.na(skew_percentile)),
    nother = dplyr::n()
  ) %>%
  ungroup()

diss_rs_results <- all_di_with_resamples %>%
  filter(nparts > 40) %>%
  group_by(Dataset, resampling) %>%
  summarize(dissimilarity_high = mean(real_po_percentile_excl > 95, na.rm = T),
            ndiss = sum(!is.na(real_po_percentile_excl))) %>%
  ungroup()

rs_results <- left_join(metric_rs_results, diss_rs_results)

#metric_cols <- colnames(rs_results [ grepl("_", colnames(rs_results))])

as.perc <- function(x) {
  return((signif(100 * x, digits = 2)))
}

>>>>

---
title: "Unusual Gentry communities"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    # github_document:
    #    df_print: kable
    #    toc: true
    word_document:
        df_print: kable
        toc: false
---

@@

@@

@@

The Gentry communities are extreme relative to hte other datasets (in grey), in that they have high species richness and low abundance. Of these, the most extreme communities have extremely low average abundance (e.g. N/S < 3).

gentry_di <- gentry_di %>%
  mutate(low_avg_abund = ifelse(n0/s0 < 3, "N:S < 3", "N:S > 3")) %>%
  mutate(`Low average \nabundance` = low_avg_abund)

ggplot(gentry_di, aes(s0, n0, color = `Low average \nabundance`)) +
  geom_point(data = all_di, inherit.aes = F, aes(s0, n0), color = "grey", alpha = .1) +  geom_point() +
scale_x_log10() + scale_y_log10() + xlab("Species richness (S); note log scale") + ylab("Total abundance (N); note log scale") + ggtitle("Gentry communities", subtitle = "(in color)")

@@

These low-average-abundance communities account for most of the signal of extreme values in the unusual direction for various metrics - that is, a **low** number of rare species and **low** skewness, and **high** Shannon and Simpson evenness. Gentry communities with average abundance > 3 tend to show the signal in the same direction as most other communities.

```{r, fig.dim = c(7,7)}
fig_2 <- gridExtra::grid.arrange(grobs = list(

@@
